### 2016-09-02 ReviewComments/response.tex

We are deeply grateful to all the reviewers for the generous and insightful
suggestions which help us to improve this work to a better level.
In the following, we provide explicit response to each raised concern and address our corresponding revisions in the new submission.

\section{Response to Reviewer 1}

\textbf{1.1} \emph{The main con that I've spotted is the fact that the algorithm might have been
more nicely framed also within the Spark environment by taking advantage of
the various possibilities offered by it (e.g. caching of RDDs)}

\textbf{Response:} We agree with the reviewer that utilizing advanced Spark features
would benefit our solutions. Indeed, our SPARE algorithm has already taken advantages of the Spark-only features: (1) caching of RDDs and (2) DAG execution engine, to achieve speedups.

As described in Section 5.3, our SPARE algorithm uses a run-time best-fit strategy to achieve load balance. The best-fit strategy allocates the largest unassigned star to the most empty reducer, which balances the workloads for each reducer. Since the strategy needs to know the sizes of all stars, it takes place after the Map phase and before the Reduce phase.
Leveraging Spark's DAG execution engine, we can easily augment traditional Map-Reduce workflow to the Map-Planning-Reduce workflow. Furthermore, we utilize the RDD caching feature of Spark to store the map result (i.e., RDDs of stars)
during the planning phase and reuse them in the reduce phase.

In Section 6.2.2 Figure 8, we validate the benefits of adopting these Spark features by comparing SPARE with
SPARE-RD. SPARE-RD is identical to SPARE except it uses random allocation rather than best-fit strategy. Benefit from the two Spark features, the SPARE the best-fit strategy
 only takes around 4\% the total execution time. In return, SPARE gains 15\% speedup compared to SPARE-RD.

### 2016-09-02 ReviewComments/meta.tex
More revisement

\textbf{Response:} In addition to the points in the response of comments 2.1, another novelty of our work lies in the 
techniques. 
Although the idea of star partition and apriori enumeration are well-known, 
linking them together to solve the problem in trajectory domain has not been attempted before.
Besides, we input heavy details on these algorithms (e.g., theoretical bounds of partition and
anti-monotonicity) which have not been previously proposed.

>>>>

\textbf{Response:} In addition to our response to comments 2.1, another novelty of our work lies in the 
techniques. 
Although the idea of star partition and apriori enumeration are well-known, 
linking them together to solve the problem in trajectory domain has not been attempted before.
Besides, we input heavy details on these algorithms (e.g., theoretical bounds of partition and
anti-monotonicity) which have not been previously proposed yet.

### 2016-09-02 ReviewComments/review2.tex
More revisement

\textbf{Response:} We admit that the gap parameter
is also used in mining Gap-restricted Sequential Patterns (GSP). 
However, our work is the first to introduce the gap concept 
in co-movement mining to resolve the
loose-connection anomaly (as shown in Figure 2).  
Moreover, we also provide
a unified model to capture existing pattern models which 
is also not proposed before.
%
%The novelty of our modeling
%is bi-folded. First, we unify existing co-movement patterns in
%a GCMP model. Second, we use the gap-constraint to avoid
%the loose-connection anomaly (as shown in Figure 2) which 
%is suffered by existing patterns.

@@

Besides, GSP and GCMP are designed with different goals. Thus it is infeasible
to model GCMP using GSP. 
In particular, the goal of GSP 
is to find the frequently occurred subsequences which conform to the gap constraint.
This definition ignores the relationships among the involved objects. 
As a consequence, the mined GSP results cannot guarantee the spatial closeness of the involved
objects, which contradicts with \emph{closeness} requirement of GCMP (Section 3 Definition 2). 
For example, in the sample trajectory database in Figure 1, 
the most frequent temporal sequence for GSP is
$(1,2,3,4,5,6)$ because it appears in all trajectories. 
However, the involved objects $\{o_1,o_2,o_3,o_4,o_5,o_6\}$ are NOT all
spatially close at any of the timestamps thus these objects do not form a proper GCMP.
We summarize this difference in Section 2.5.

>>>>

\textbf{Response:} We admit that the gap parameter
is also used in mining Gap-restricted Sequential Patterns (GSP). 
However, our work is the first to introduce the gap constraint 
in co-movement mining in trajectory domain, and we are the first to resolve the
loose-connection anomaly (as shown in Figure 2).  
Moreover, we provide
a unified model to capture existing pattern models which 
is also not proposed before.
%
%The novelty of our modeling
%is bi-folded. First, we unify existing co-movement patterns in
%a GCMP model. Second, we use the gap-constraint to avoid
%the loose-connection anomaly (as shown in Figure 2) which 
%is suffered by existing patterns.

@@

Besides, GSP and GCMP are designed with different goals. Thus it is infeasible
to model GCMP using GSP. 
In particular, the goal of GSP 
is to find the frequently occurred subsequences which conform to the gap constraint.
This definition ignores the relationships among the involved objects. 
As a consequence, the mined GSP results cannot guarantee the spatial closeness of the involved
objects, which contradicts with \emph{closeness} requirement of GCMP (Section 3 Definition 2). 
For example, in the sample trajectory database in Figure 1, 
the most frequent temporal sequence for GSP is
$(1,2,3,4,5,6)$ because it appears in every trajectory. 
However, the involved objects $\{o_1,o_2,o_3,o_4,o_5,o_6\}$ are NOT all
spatially close at any of the timestamps thus these objects do not form a proper GCMP.
We summarize this difference in Section 2.5.

### 2016-09-02 ReviewComments/review3.tex
More revisement

\textbf{Response:}
We thank the reviewer for suggesting a better way of implementing 
the baseline algorithm (i.e., TRPM). By utilizing window functions,
TRPM could be rewritten as in the following pseudo-code:
\begin{equation*}
\begin{aligned}
 \mathtt{WindowSpec}\ ws := \mathtt{Window}&.\mathtt{orderBy}('time') \\
 				&.\mathtt{rangeBetween}(0, \eta); \\
 \mathtt{Pattern}\ pt :=  LineSweep&('clusters').\mathtt{over}(ws); 
\end{aligned}
\end{equation*}

@@

On the other hand, we believe the performance boost on TRPM from Spark-SQL
would be very limited. The reason is that our 
Line Sweep algorithm (Algorithm 1) is not a 
distributive function. 
That is the result over an input cannot be constructed from
the result over a subset of the input. This indicates 
that no partial results can be shared among all windows.
In such a case, Spark-SQL has to process each
window independently in parallel, which is equivalent to our TRPM implementation.
%We agree that leveraging more advanced Spark features
%could further improve the performance of our solutions. 
%We do not heavily leverage these features because 
%we wish to focus on designing parallel GCMP mining algorithms.
%%to provide a general parallel framework which does not tie to Spark.
%Nevertheless, we are in progress in developing
%a better open-sourced version with Spark-tailored optimization. 
%In the revision, we try to add a DataFrame based TRPM solution 
%which aims to leverage the window functions from Spark-SQL.
%However, to the best of our efforts, we could not achieve this goal in a short time. 
%The major challenge is that current Spark do not support User Defined Aggregate Function (UDAF)
%in window functions~\footnote{JIRA Spakr-8641: Native Spark Window Functions \url{https://issues.apache.org/jira/browse/SPARK-8641}}. Without UDAF, implementing
%our Line sweep algorithm (Algorithm 1) using Spark primitives (e.g., sum, avg, rank and nth-tile)
%is nontrivial. Nonetheless, we expect less performance
%boost from TRPM even with the UDAF support. This is because the Line sweep algorithm is not
%properly reducible, where Spark system would fail to leverage the partial aggregates
%across multiple windows. As a result, it would be equivalent to our current TRPM
%solution where windows are processed independently in parallel.
%TRPM although can be modeled using window functions over data frame, the aggregates
%it uses (i.e., the Line Sweeping method in Algorithm 1) cannot be simply represented
%by Spark primitives (e.g., sum, avg, rank, nth-tile). We then resort to
%the Spark User Defined Aggregate Function (UDAF). However, in Spark 1.5.2, the UDAF
%does not support window function syntax. Even in the latest Spark 1.6.0 (released 
%last month), UDAF does not fit with window function. We confirm these information
%with Spark development team~\footnote{In Spark JIRA tickets, fully supporting UDAF is still under development}.
%
%Nevertheless, since our Algorithm 1 is not reducible, even with UDAF, TRPM could
%not boost too much 
%
%we expect
%even with UDAF support TRPM would not boost too much.  Spark system is 
%%hard to leverage the partial aggregates acrocess
%
%we can analytically expect even using UDAF, TRPM would
%not boost too much. This is because our Algorithm 1 is not reducible. That
%is we cannot merge or share partial aggregates across different windows. In summary,
%We wish to convince the reviewer that current TRPM implementation is reasonable.

@@

\textbf{Response:}  We study the skewness by examining the load balance of each algorithm.
We revise Section 6.2.2 by comparing SPARE with different data partition strategies.
In particular, we compare SPARE which adopts a best-fit strategy (Section 5.3) and SPARE-RD
which adopts the default random partition strategy.
We present the summary of the execution time for each method on the longest job, the shortest job and standard
deviation of all jobs in Table 9. As the table shows, SPARE has a better load balance (low deviation)
than SPARE-RD which indicates the effectiveness of the best-fit strategy. 
%Further, TRPM generally keeps a low deviation. This is because that TRPM always adopts the equal-size (i.e. $\eta$) partition. Nevertheless, both the longest and shortest jobs in TRPM are higher than SPARE based solutions.

>>>>

\textbf{Response:}
We thank the reviewer for suggesting a better way to implement
the baseline algorithm (i.e., TRPM). With window functions,
TRPM could be rewritten as in the following pseudocode:
\begin{equation*}
\begin{aligned}
 \mathtt{WindowSpec}\ ws := \mathtt{Window}&.\mathtt{orderBy}('time') \\
 				&.\mathtt{rowBetween}(0, \eta); \\
 \mathtt{Pattern}\ pt :=  LineSweep&('clusters').\mathtt{over}(ws); 
\end{aligned}
\end{equation*}

@@

On the other hand, we believe the performance boost on TRPM from Spark-SQL
would be very limited. The reason is that our 
Line Sweep algorithm (Algorithm 1) is
distributive.
That is the result over an input cannot be constructed from
the result over a subset of the input. This indicates 
that no partial results could be shared among all windows.
In such a case, Spark-SQL has to process each
window independently in parallel, which is equivalent to our TRPM implementation.
%We agree that leveraging more advanced Spark features
%could further improve the performance of our solutions. 
%We do not heavily leverage these features because 
%we wish to focus on designing parallel GCMP mining algorithms.
%%to provide a general parallel framework which does not tie to Spark.
%Nevertheless, we are in progress in developing
%a better open-sourced version with Spark-tailored optimization. 
%In the revision, we try to add a DataFrame based TRPM solution 
%which aims to leverage the window functions from Spark-SQL.
%However, to the best of our efforts, we could not achieve this goal in a short time. 
%The major challenge is that current Spark do not support User Defined Aggregate Function (UDAF)
%in window functions~\footnote{JIRA Spakr-8641: Native Spark Window Functions \url{https://issues.apache.org/jira/browse/SPARK-8641}}. Without UDAF, implementing
%our Line sweep algorithm (Algorithm 1) using Spark primitives (e.g., sum, avg, rank and nth-tile)
%is nontrivial. Nonetheless, we expect less performance
%boost from TRPM even with the UDAF support. This is because the Line sweep algorithm is not
%properly reducible, where Spark system would fail to leverage the partial aggregates
%across multiple windows. As a result, it would be equivalent to our current TRPM
%solution where windows are processed independently in parallel.
%TRPM although can be modeled using window functions over data frame, the aggregates
%it uses (i.e., the Line Sweeping method in Algorithm 1) cannot be simply represented
%by Spark primitives (e.g., sum, avg, rank, nth-tile). We then resort to
%the Spark User Defined Aggregate Function (UDAF). However, in Spark 1.5.2, the UDAF
%does not support window function syntax. Even in the latest Spark 1.6.0 (released 
%last month), UDAF does not fit with window function. We confirm these information
%with Spark development team~\footnote{In Spark JIRA tickets, fully supporting UDAF is still under development}.
%
%Nevertheless, since our Algorithm 1 is not reducible, even with UDAF, TRPM could
%not boost too much 
%
%we expect
%even with UDAF support TRPM would not boost too much.  Spark system is 
%%hard to leverage the partial aggregates acrocess
%
%we can analytically expect even using UDAF, TRPM would
%not boost too much. This is because our Algorithm 1 is not reducible. That
%is we cannot merge or share partial aggregates across different windows. In summary,
%We wish to convince the reviewer that current TRPM implementation is reasonable.

@@

\textbf{Response:}  We study the skewness by examining the load balance of each algorithm.
We revise Section 6.2.2 by comparing SPARE with different data partition strategies.
In particular, we compare SPARE which contains a best-fit strategy (Section 5.3) and SPARE-RD
which adopts the default random partition strategy.
We present the summary of the execution time for each method on the longest job, the shortest job and standard
deviation of all jobs in Table 9. As the table shows, SPARE has a better load balance (low deviation)
than SPARE-RD which indicates the effectiveness of the best-fit strategy. 
%Further, TRPM generally keeps a low deviation. This is because that TRPM always adopts the equal-size (i.e. $\eta$) partition. Nevertheless, both the longest and shortest jobs in TRPM are higher than SPARE based solutions.

### 2016-09-02 sec6_experiment.tex
More revisement

\begin{table}[h]
\centering \small
%\begin{tabular}{|c|c|c|c|c|}
%\hline
%\multirow{2}{*}{\textbf{Dataset}} & \multicolumn{2}{c|}{\textbf{SPARE-RD}} & \multicolumn{2}{c|}{\textbf{SPARE}} \\ \cline{2-5} 
%                         & Straggler        & Std. Dev.     & Straggler       & Std. Dev.      \\ \hline
%Shopping                 & 295              & 41         & 237             & 21       \\ \hline
%GeoLife                  & 484              & 108        & 341             & 56       \\ \hline
%Taxi                     & 681              & 147        & 580             & 96       \\ \hline
%\end{tabular}
\begin{tabular}{|l|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{\textbf{Dataset}} & \textbf{Method} & \textbf{Min} & \textbf{Max} & \textbf{Std. Dev.} 
\\ \hline%\specialrule{1pt}{-0.5pt}{-0.5pt} 
\multirow{3}{*}{Shopping}              & TRPM            &     426                          & 597                 & 35 
\\ %\cline{2-5} 
                                       & SPARE-RD          & 122                               &  295             & 41
\\% \cline{2-5}                                        
									   & SPARE           &  134                              &   237               & 21
\\ \hline                                
\multirow{3}{*}{GeoLife}              & TRPM            &    493                            & 747                & 78               \\ %\cline{2-5} 
                                       & SPARE-RD          &  38                               &   484               & 148
\\ %\cline{2-5}                                        
									   & SPARE           &   107                             &     341             & 96
\\ \hline
\multirow{3}{*}{Taxi}              & TRPM            &   3,460                            &    3,648              & 51 \\ %\cline{2-5} 
                                       & SPARE-RD          & 122                               & 681                 & 109
\\ %\cline{2-5}                                        
									   & SPARE           &   134                             &  580                & 56               
\\ \hline
\end{tabular}
  \caption{Statistics of execution time (in seconds) among all executors.}
  \label{tbl:strags}
\end{table}

\subsubsection{Load balance}
To study the effect of load balance in the SPARE framework, we use random task allocation (the default setting of Spark) as a baseline, denoted by SPARE-RD, and compare it with our best-fit method. In best-fit, the largest unassigned star is allocated to the currently most lightly loaded reducer.
Figure~\ref{exp:wl} shows the breakdown of the costs in the map-reduce stages 
for SPARE and SPARE-RD \revised{with TRPM as comparison}. We observe that the map and shuffle time of SPARE and SPARE-RD are identical . The difference is that SPARE incurs an additional overhead to generate an allocation plan for load balance (around $4\%$ of the total cost), resulting in significant savings in the reduce stage (around $20\%$ of the total cost). 
\revised{Meanwhile, both SPARE and SPARE-RD outperform TRPM in each phase. This indicates the 
efficiency of the star partition and apriori enumeration.} 
\revised{We also report the cost summaries (i.e., longest, shortest and standard deviation) of all jobs for the three methods in Table~\ref{tbl:strags}. As the table shows, SPARE produces more balanced jobs
than SPARE-RD which verify the effectiveness of the best-fit strategy. In addition, TRPM 
maintains a low deviation because it uses equal size (i.e., $\eta$) partitioning. 
However, TRPM still takes much longer time than SPARE algorithms.}
%We also report the cost  for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.
%We also report the cost of \emph{straggler}, i.e., the longest job, and the standard deviation (Std. Dev.) for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.

>>>>

\begin{table}[h]
\centering \small
%\begin{tabular}{|c|c|c|c|c|}
%\hline
%\multirow{2}{*}{\textbf{Dataset}} & \multicolumn{2}{c|}{\textbf{SPARE-RD}} & \multicolumn{2}{c|}{\textbf{SPARE}} \\ \cline{2-5} 
%                         & Straggler        & Std. Dev.     & Straggler       & Std. Dev.      \\ \hline
%Shopping                 & 295              & 41         & 237             & 21       \\ \hline
%GeoLife                  & 484              & 108        & 341             & 56       \\ \hline
%Taxi                     & 681              & 147        & 580             & 96       \\ \hline
%\end{tabular}
\begin{tabular}{|l|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{\textbf{Dataset}} & \textbf{Method} & \textbf{Min} & \textbf{Max} & \textbf{Std. Dev.} 
\\ \hline%\specialrule{1pt}{-0.5pt}{-0.5pt} 
\multirow{3}{*}{Shopping}             % & TRPM            &     426                          & 597                 & 35 \\ %\cline{2-5} 
                                       & SPARE-RD          & 122                               &  295             & 41
\\% \cline{2-5}                                        
									   & SPARE           &  134                              &   237               & 21
\\ \hline                                
\multirow{3}{*}{GeoLife}             % & TRPM            &    493                            & 747                & 78               \\ %\cline{2-5} 
                                       & SPARE-RD          &  38                               &   484               & 148
\\ %\cline{2-5}                                        
									   & SPARE           &   107                             &     341             & 96
\\ \hline
\multirow{3}{*}{Taxi}             % & TRPM            &   3,460                            &    3,648              & 51 \\ %\cline{2-5} 
                                       & SPARE-RD          & 122                               & 681                 & 109
\\ %\cline{2-5}                                        
									   & SPARE           &   134                             &  580                & 56               
\\ \hline
\end{tabular}
  \caption{Statistics of execution time (in seconds) among all executors.}
  \label{tbl:strags}
\end{table}

\subsubsection{Load balance}
To study the effect of load balance in the SPARE framework, we use random task allocation (the default setting of Spark) as a baseline, denoted by SPARE-RD, and compare it with our best-fit method. In best-fit, the largest unassigned star is allocated to the currently most lightly loaded reducer.
Figure~\ref{exp:wl} shows the breakdown of the costs in the map-reduce stages 
for SPARE and SPARE-RD. %\revised{with TRPM as comparison}. 
We observe that the map and shuffle time of SPARE and SPARE-RD are identical . The difference is that SPARE incurs an additional overhead to generate an allocation plan for load balance (around $4\%$ of the total cost), resulting in significant savings in the reduce stage (around $20\%$ of the total cost). 
\revised{Meanwhile, both SPARE and SPARE-RD outperform TRPM in each phase. This indicates the 
efficiency of the star partition and apriori enumeration.} 
%\revised{We also report the cost summaries (i.e., longest, shortest and standard deviation) of all jobs for the three methods in Table~\ref{tbl:strags}. As the table shows, SPARE produces more balanced jobs
%than SPARE-RD which verify the effectiveness of the best-fit strategy. In addition, TRPM 
%maintains a low deviation because it uses equal size (i.e., $\eta$) partitioning. 
%However, TRPM still takes much longer time than SPARE algorithms.}
We also report the cost  for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.
We also report the cost of \revised{the longest job (Max), the shortest job (Min) and the standard deviation (Std. Dev.)}
%\emph{straggler}, i.e., the longest job, and the standard deviation (Std. Dev.) 
for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.
